### 2018-05-04 papers/synth/intro.tex
somya added

Eukaryotic cells, including human cells, consist of multiple membrane bound compartments.
Material is transported among these compartments by the vesicle transport system (VTS) []. Briefly,
the source compartment produces a membrane bound packet of molecules called a vesicle. After
release, this vesicle specifically recognizes the correct target compartment within the cell, and fuses
with it. The molecules that form the machinery of the VTS are known. \todo{connect to function }{We also know a lot about the specific functions} of these molecules and how these molecules interact amongst each other. In spite of this detailed knowledge at the level of the molecules, the structure of the VTS network, or the road-map of the eukaryotic cell, is far from complete. 
%The current knowledge of the network is put together from a patch work of biological experiments. This information is scattered across several publications. Even after this information is collected and put together, we find that the network obtained is still not complete; new vesicles and new contents in previously known vesicles are constantly being discovered. The synthesis for the unknown pieces may be assisted by computationon the graph model of VTSs. 
% 
%In this work, we use a model described in [] to find out, given a partial
%VTS graph and details about interactions amongst its molecules, what the complete graph of the
%VTS could look like. (Please look at appendix A for the detailed discussion on pros and cons of the
%model.
%
%Eukaryotic cells, including human cells, consists of multiple compartments.
%
%Vesicle Traffic Systems(VTSs) are the material transport mechanisms among the compartments inside the cells~\cite{vtsIntro}.
%
%Almost all subsystems of the cells depend on VTSs.
%
%Therefore, understanding how VTSs functions is one of the key
%questions of cell biology.
%
In this paper, we are looking at the computational questions 
arise from the VTSs.
%
In the following, we use the model of VTSs that has been presented
in~\cite{VTS}.
%
Please look at appendix~\ref{sec:model}
for the detailed discussion on pros and cons of the model.

>>>>

Eukaryotic cells, including human cells, consist of multiple membrane bound compartments.
Material is transported among these compartments by the vesicle transport system (VTS) []. Briefly,
the source compartment produces a membrane bound packet of molecules called a vesicle. After
release, this vesicle specifically recognizes the correct target compartment within the cell, and fuses
with it. The molecules that form the machinery of the VTS are known. \todo{connect to function }{We also know a lot about the specific functions} of these molecules and how these molecules interact amongst each other. In spite of this detailed knowledge at the level of the molecules, the structure of the VTS network, or the road-map of the eukaryotic cell, is far from complete. The current knowledge of the network is put together from a patch work of biological experiments. This information is scattered across several publications. Even after this information is collected and put together, we find that the network obtained is still not complete; new vesicles and new contents in previously known vesicles are constantly being discovered. The synthesis for the unknown pieces may be assisted by computation on the graph model of VTSs. 
% 
%In this work, we use a model described in [] to find out, given a partial
%VTS graph and details about interactions amongst its molecules, what the complete graph of the
%VTS could look like. (Please look at appendix A for the detailed discussion on pros and cons of the
%model.
%
%Eukaryotic cells, including human cells, consists of multiple compartments.
%
%Vesicle Traffic Systems(VTSs) are the material transport mechanisms among the compartments inside the cells~\cite{vtsIntro}.
%
%Almost all subsystems of the cells depend on VTSs.
%
%Therefore, understanding how VTSs functions is one of the key
%questions of cell biology.
%
In this paper, we are looking at the computational questions 
arise from the VTSs.
%
In the following, we use the model of VTSs that has been presented
in~\cite{VTS}.
%
Please look at appendix~\ref{sec:model}
for the detailed discussion on pros and cons of the model.

### 2018-05-04 papers/synth/model.tex
somya added

\ankit{Write about Q/R SNARES: fill tomorrow}
The molecules transported by the VTS are themselves its regulators. The molecules in a
compartment/vesicle may be active or inactive. The molecules that are respnsible for vesicle fusion
are called SNARE proteins []. Active SNAREs present on vesicles (v-SNAREs) bind with their
cognate active SNAREs on the target compartment (t-SNAREs) to enable veiscle fusion. A cell
contains multiple kinds of v- and t- SNAREs. Only specific pairs of v and t SNAREs can bind to
each other and participate in fusion. Fusion compatible v- and t- SNAREs are determined by
biological experiments. Different vesicle-compartmetn fusions in the cell are brought about by
different v- and t-SNARE pairs. A molecule that participates in a given fusion reaction must not
interfere with fusion at different compartments or vesicles. Therefore, SNAREs must be kept in an
inactive form in appropriate compartments/vesicles. The activity of molecules is regulated by the
other molecules, i.e., the presence and absence of the other molecules in a compartment or vesicle
may make the molecule active or inactive. We call this regulation as activity functions.
In the VTS model, we assume that the system is in steady state and the concentrations of the
molecules in the compartments do not change over time. We define SNARE pairing specificity by a
fusion pairing relation containing pairs of SNAREs and molecular regulation by activity Boolean
functions. Since the system is in steady state, we expect that any molecule that leaves a
compartment must come back via some path on the graph. We call this property of VTS as stability.

\ankit{Discuss the yeast and eukaryotes VTS model and assumptions }
A few assumptions on biological level.

\begin{enumerate}
\item[1] A cell is a set of compartments exchanging vesicles.
\item[2] Compartments are neither created nor destroyed.
\item[3] Each compartment is in steady state, gain and loss balance.
\item[4] Molecules are neither created nor destroyed.
\item[5] Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. \ankit{Change it to consider QR SNARES}
 \end{enumerate}

\begin{enumerate}

\item Complex form using three Q-SNAREs and one R-SNARE. In order for a vesicle to fuse to it's corresponding target compartment this combination has to be present.
\item In the graph shown the R-SNAREs are always in active. Every combination is formed by three active Q-SNAREs and one active R-SNARE.
\item  Qai, Qbi, Qci and Ri are just dummy names, real biological names can be found in the below table. 
\item In the diagram white circle molecules are always active. Other edge labels are inactive molecules.
\item Pairing matrix restriction: ith R-SNARE can fuse with only the corresponding ith Q-SNARES and vice-versa. No other combination is allowed.
\item Concept of v and t-SNAREs is replaced by Q and R-SNAREs. 
\end{enumerate}

>>>>

The molecules transported by the VTS are themselves its regulators. The molecules in a
compartment/vesicle may be active or inactive. The molecules that are respnsible for vesicle fusion
are called SNARE proteins []. Active SNAREs present on vesicles (v-SNAREs) bind with their
cognate active SNAREs on the target compartment (t-SNAREs) to enable vesicle fusion. A cell
contains multiple kinds of v- and t-SNAREs. Only specific pairs of v and t SNAREs can bind to
each other and participate in fusion. Fusion compatible v- and t- SNAREs are determined by
biological experiments. Different vesicle-compartment fusions in the cell are brought about by
different v- and t-SNARE pairs. A molecule that participates in a given fusion reaction must not
interfere with fusion at different compartments or vesicles. Therefore, SNAREs must be kept in an
inactive form in appropriate compartments/vesicles. The activity of molecules is regulated by the
other molecules, i.e., the presence and absence of the other molecules in a compartment or vesicle
may make the molecule active or inactive. We call this regulation as activity functions.
In the VTS model, we assume that the system is in steady state and the concentrations of the
molecules in the compartments do not change over time. We define SNARE pairing specificity by a
fusion pairing relation containing pairs of SNAREs and molecular regulation by activity Boolean
functions. Since the system is in steady state, we expect that any molecule that leaves a
compartment must come back via some path on the graph. We call this property of VTS as stability.

A few assumptions on modeling level.
\begin{enumerate}
\item[1] A cell is a set of compartments exchanging vesicles.
\item[2] Compartments are neither created nor destroyed.
\item[3] Each compartment is in steady state, gain and loss balance.
\item[4] Molecules are neither created nor destroyed.
\item[5] Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}

\ankit{Discuss the yeast and eukaryotes VTS model and assumptions }

\begin{enumerate}
\item Complex form using three Q-SNAREs and one R-SNARE. In order for a vesicle to fuse to it's corresponding target compartment this combination has to be present.
\item In the graph shown the R-SNAREs are always in active. Every combination is formed by three active Q-SNAREs and one active R-SNARE.
\item  Qai, Qbi, Qci and Ri are just dummy names, real biological names can be found in the below table. 
\item In the diagram white circle molecules are always active. Other edge labels are inactive molecules.
\item Pairing matrix restriction: ith R-SNARE can fuse with only the corresponding ith Q-SNARES and vice-versa. No other combination is allowed.
\end{enumerate}

### 2018-05-04 papers/synth/related.tex
somya added

There are a few examples of using synthesis technique in biology especially in gene regulatory networks~\cite{shavit2016automated, fisher2015synthesising}. A very recent work~\cite{fisher2015synthesising} synthesize executable gene regulatory networks from single-cell gene expression data. 
Synthesis technique is also used in optimal synthesis for chemical reaction networks~\cite{cardelli2017syntax}. the~\cite{fisher2015synthesising} uses constraint (satisfiability) solving techniques for the synthesis whereas ~\cite{shavit2016automated} uses SMT for synthesis. The queries contain quantifiers so we have we employ QBF solving for the solving the synthesis problem. The paper~\cite{cardelli2017syntax} uses a template-guided approach which is very close to our work. To our best knowledge this is the first application of synthesis in VTS.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

There are a few examples of using synthesis technique in biology especially in gene regulatory networks~\cite{shavit2016automated, fisher2015synthesising}. A very recent work~\cite{fisher2015synthesising} synthesize executable gene regulatory networks from single-cell gene expression data. 
Synthesis technique is also used in optimal synthesis for chemical reaction networks~\cite{cardelli2017syntax}. The~\cite{fisher2015synthesising} uses constraint (satisfiability) solving techniques for the synthesis whereas ~\cite{shavit2016automated} uses SMT for synthesis. The paper~\cite{cardelli2017syntax} in addition of using SMT over ODE, uses a template-guided approach. In our case queries contain quantifiers so we have employed QBF solving with Z3 for the solving the synthesis problem. To our best knowledge this is the first application of synthesis in VTS.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
